---
title: "Coursera test 2"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---


```{r}
train.data <- readRDS(file="training_data_10.RDS")
source("main.R")
```



## 1 
When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd

- sleep 
- die +
- give
- eat

Posible Response:

```{r}

query<-"When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```






## 2 
Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his

- financial 
- spiritual
- marital + 
- horticultural

Posible Response:

```{r}

query<-"Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```



## 3 
I'd give anything to see arctic monkeys this

- month  + 
- morning
- decade
- weekend

Posible Response:

```{r}

query<-"I'd give anything to see arctic monkeys this"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 4 
Talking to your mom has the same effect as a hug and helps reduce your

- stress + 
- happiness
- sleepiness
- hunger

Posible Response:

```{r}

query<-"Talking to your mom has the same effect as a hug and helps reduce your"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 5 
When you were in Holland you were like 1 inch away from me but you hadn't time to take a

- look ++++++++ 
- picture
- minute
- walk

Posible Response:

```{r}

query<-"When you were in Holland you were like 1 inch away from me but you hadn't time to take a"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 6 
I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the

- case + 
- matter
- incident
- account

Posible Response:

```{r}

query<-"I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 7
I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each

- hand 
- finger
- toe
- arm + 

Posible Response:

```{r}

query<-"I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 8
Every inch of you is perfect from the bottom to the

- side 
- middle
- center
- top + 

Posible Response:

```{r}

query<-"Every inch of you is perfect from the bottom to the"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


## 9
I’m thankful my childhood was filled with imagination and bruises from playing

- outside + 
- inside
- weekly
- daily

Posible Response:

```{r}

query<-"I’m thankful my childhood was filled with imagination and bruises from playing"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```



## 10 
I like how the same people are in almost all of Adam Sandler's

- pictures
- stories
- movies + 
- novels

Posible Response:

```{r}

query<-"I like how the same people are in almost all of Adam Sandler's"
solution <-predict.word (query) 
solution$candidates[,c("full.sentence","words","prob")]

```


