---
title: 'Text Prediction: Exploratory data analysis'
author: "Enrique Repulles"
output:
  pdf_document:
    number_sections: yes
  html_notebook:
    number_sections: yes
  html_document:
    number_sections: yes
---


```{r imports, , include=FALSE}
# The libraries needed to run the code
library("NLP")
library("tm")
library("ggplot2")
library("knitr")
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.align="center", cache = TRUE)

```



Load data and save in RDS format

```{r, eval=FALSE, include=FALSE}
sampleFile<-function(myfilename, sample.proportion)
{
  path<-paste0("./raw_data/final/en_US/",myfilename)
  out.path<-paste0("./clean_data/",strsplit(myfilename,split = ".txt")[[1]],".sampled.txt")
print(out.path)  
  raw.data = readLines(path,warn = FALSE,skipNul = TRUE);
  sample<-1==rbinom(n=length(raw.data), p=sample.proportion,size=1)
  data<-raw.data[sample]
  # saveRDS(object = data, file = out.path)
  writeLines(data,con=out.path)
  
}

# sampleFile("en_US.blogs.txt", .001)
# sampleFile("en_US.news.txt", .001)
# sampleFile("en_US.twitter.txt", .001)


```




```{r}
?VectorSource

corpus.source<-DirSource("./clean_data/")
corpus.source
corpus<-SimpleCorpus(corpus.source,control = list(language = "en"))
summary(corpus)


```

```{r}
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus , stripWhitespace)
corpus <- tm_map(corpus, tolower)
corpus <- tm_map(corpus, removeWords, stopwords("english")) # quitar
```




```{r}

tdm <- TermDocumentMatrix(corpus)
inspect(tdm)




```




```{r}
dtm <- DocumentTermMatrix(corpus)
inspect(dtm)
```


```{r}
#install.packages("RWeka")
#install.packages("rJava")

library(RWeka)

?NGramTokenizer
```




```{r}
TrigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 3, max = 3))
tdm3 <- TermDocumentMatrix(corpus, control = list(tokenizeNO_FUCIONA = TrigramTokenizer))
tdm3 <- removeSparseTerms(tdm3, 0.75)
inspect(tdm3)

```




- tm::c.VCorpus		Combine Corpora, Documents, Term-Document Matrices, and Term Frequency Vectors
- tm::findAssocs		Find Associations in a Term-Document Matrix
- tm::findMostFreqTerms		Find Most Frequent Terms
- tm::inspect		Inspect Objects
- tm::TermDocumentMatrix		Term-Document Matrix
- tm::plot.TermDocumentMatrix		Visualize a Term-Document Matrix
- tm::tm_term_score		Compute Score for Matching Terms
- tm::plot.TermDocumentMatrix		Visualize a Term-Document Matrix


