---
title: 'Text Prediction: Exploratory data analysis'
author: "Enrique Repullés"
output:
  html_notebook:
    number_sections: yes
  html_document:
    number_sections: yes
  pdf_document:
    number_sections: yes
---


```{r imports, , include=FALSE}
# The libraries needed to run the code
library("NLP")
library("tm")
library("ggplot2")
library("knitr")
```


```{r global_options, include=FALSE}
#knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.align="center", cache = TRUE)

```

# Introduction

The purpose of this document is to show the exploratory analysis and text mining done in the text corpus and to preview some future directions in the construction of the text prediction algorithm.

The source for this document and other project code can be found in [github.com/enrique-repulles/predictive-text](https://github.com/enrique-repulles/predictive-text).

```{r preliminaries}
## Preliminaries 

set.seed(1234)
#sample.size=.01
sample.size=.001
```


# Data sources

The text corpus was obtained from  

[Coursera-SwiftKey.zip](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip)

An script for download this file can be found in the project github page.

For the "profanity filtering" cleaning, we've used an offensive words list from  [Terms-to-Block.csv](http://www.frontgatemedia.com/new/wp-content/uploads/2014/03/Terms-to-Block.csv)



# Data cleaning and preprocessing


The first step was to extract a sample of the corpus (a `r sample.size*100`%) and to split the corpus into sentences. 
For deciding where to split the text, we chose some characters to mark the end of a sentence:'.',';'...

After that, we removed weird symbols and unified the format (lower case , single spaces between words...).  Also, any offensive word ("profanity filtering") are removed.

About stop words (like 'the', 'a', 'an'...): we've decided to keep them. In text classification they are usually removed but our guess is  that for prediction they will be useful.



```{r tokenization_functions}

# Read the list of offensive words 
offensive.words<-readLines("raw_data/Terms-to-Block.csv")
offensive.words<-offensive.words[5:length(offensive.words)]
offensive.words<-gsub("[\"]","",offensive.words)
offensive.words<-gsub("[,]","",offensive.words)


# The *loadFile* function loads the file and extract a random subset of lines.
loadFile<-function(myfilename, sample.proportion)
{
  path<-paste0("./raw_data/final/en_US/",myfilename)
  raw.data = readLines(path,warn = FALSE,skipNul = TRUE);
  sample<-1==rbinom(n=length(raw.data), p=sample.proportion,size=1)
  raw.data[sample]
}

extractSentences<-function (text)
{
  endsentence.symbols <- "[.;()!?:]"
  sentences<-unlist(strsplit(text, split=endsentence.symbols))
  sentences<-sentences[nchar(sentences)>5]
}



cleanSentences <-function(rawSentences)
{
  cleanSentences<-rawSentences
  cleanSentences<-tolower(cleanSentences)
  cleanSentences<-stripWhitespace(cleanSentences)
  cleanSentences<-removeNumbers(cleanSentences)
  cleanSentences<-removePunctuation(cleanSentences, preserve_intra_word_dashes = TRUE)
  # english_stop_words<-stopwords()
  # cleanSentences<-removeWords(cleanSentences,english_stop_words)
  cleanSentences<-removeWords(cleanSentences,offensive.words)
  cleanSentences<-trimws(stripWhitespace(cleanSentences))
}




# We include this function for testing purposes
testCleaning <- function(sentences)
{
  data.frame(raw=sentences, clean=cleanSentences(sentences))
}



# This function get a sentence and return a list of tokens. The "ngram.size" parameter indicates the size of the tokens, so for ngram.size=1, the function returns a list of words, for ngram.size=2 returns a list of bigrams, ngram.size=3 returns a list of trigrams, etc 


tokenize <- function (sentence, ngram.size)
{
  words <- unlist(strsplit(sentence," "))
  result.size <- length(words) - ngram.size + 1 
  result<- character(result.size)
  for(i in 1:result.size)
  {
    ngram <- paste(words[i:(i+ngram.size-1)], collapse=" ")
    result[i]<-ngram
  }
  result
}


# File tokenization
# The *tokenizeFile* process a whole file: loads the file, performs cleaning and tokenization and returns a list of tokens.
# The tokens returned are ngrams of several sizes, from 1-grams (words) to 4-grams.
# The result is organize in a data frame, whith one row for each token and the following fields: 
# - filename: the filethe token cames from.
# - ngram.size: 1,2,3,4 indicating if the token is a 1-gram(single word), 2-gram, 3-gram, 4-gram
# - token:  the text of the token


# File to extract counts and statistics about a file 
countFileData <- function(filename)
{
  text<-loadFile(filename, 1)
  totalrows<-length(text)
  text<-loadFile(filename, sample.size)
  samplerows<-length(text)

  rawSentences<-extractSentences(text)
  samplesentences<-length(rawSentences)
  
  data.frame(filename=filename, totalrows,samplerows,samplesentences )
}



tokenizeFile <- function(filename)
{
  text<-loadFile(filename, sample.size)
  rawSentences<-extractSentences(text)
  cleanSentences<-cleanSentences(rawSentences)
  
  ngrams1<-data.frame(token=tokenize(cleanSentences,1)) # Single words
  ngrams2<-data.frame(token=tokenize(cleanSentences,2)) # Bigrams
  ngrams3<-data.frame(token=tokenize(cleanSentences,3)) # trigrams
  ngrams4<-data.frame(token=tokenize(cleanSentences,4)) # 4-grams
  
  # Organize results in a single dataframe
  
  ngrams1$ngram.size<-1
  ngrams2$ngram.size<-2
  ngrams3$ngram.size<-3
  ngrams4$ngram.size<-4
  
  result <- rbind(ngrams1,ngrams2,ngrams3,ngrams4)
  result$origin<-filename
  result
}


# Using the functions described in the previous section, we process the three 
# files from the corpus. The result is stored in a file for the following analysis step.

# Basic counts for statistics


fileCounts<-rbind(countFileData("en_US.blogs.txt"),
                  countFileData("en_US.news.txt"),
                  countFileData("en_US.twitter.txt"))

```


Below some counts of the corpus files are shown:


```{r show_file_counts}
kable(fileCounts, col.names = c("Filename","Total rows","rows in sample","sentences in sample"), caption="Corpus counts")

```


```{r plot_file_counts}

ggplot(fileCounts, aes(x=filename,weight=totalrows)) + geom_bar()

```



The following plot shows a comparison between the number of rows and the number of sentences in the sampled text.


```{r plot_sample_rows_sentences}


plotdata1<-data.frame(filename=fileCounts$filename, weight=fileCounts$samplerows) 
plotdata1$type <- "sample rows"
plotdata2<-data.frame(filename=fileCounts$filename, weight=fileCounts$samplesentences) 
plotdata2$type <- "sample sentences"

ggplot(rbind(plotdata1,plotdata2), aes(x=filename,weight=weight, fill=type)) + geom_bar(position="dodge")

```



## Word frequency analysis

In this section we analyze the frequency of the words in the corpus. Different kinds of ngrams are also analyzed. 

From now on, the concept of "ngram size" will be used, meaning the number of words in a token. So, ngram size=1 means a single word, ngram size=2 are bigrams, 3 trigram a 4 quadrigrams. 

```{r corpus_tokenization}

# Process execution

file1Processing<-tokenizeFile("en_US.blogs.txt")
file2Processing<-tokenizeFile("en_US.news.txt")
file3Processing<-tokenizeFile("en_US.twitter.txt")

tokens<-rbind(file1Processing,file2Processing,file3Processing)
tokens$token <- as.character(tokens$token)

#cleanup
rm(file1Processing,file2Processing,file3Processing)

wordCounts<-aggregate(tokens$ngram.size, by=list(filename=tokens$origin, ngram.size=tokens$ngram.size), length)
```



Below we shown a summary of word counts grouped by file and ngram size: 



```{r show_word_counts}
# word counts

kable(wordCounts, col.names = c("Filename","ngram size","count"), caption = "Token counts by file and ngram size")

```


And a plot with the same information: 


```{r plot_word_counts}
ggplot(wordCounts, aes(x=filename,weight=x)) + geom_bar()
```



## Word frequency distribution 

Now we examine the distributions of the frequencies, that is how many words appears one time, how many appears two times etc. 
In this sections, the data of the three files are add together.


```{r process_frequencies}

# calculations 

frequency<-function(ngram.size)
{
  
  df<-as.data.frame(table(tokens[tokens$ngram.size==ngram.size,]$token))
  df$ngram.size<-ngram.size
  data.frame(ngram.size=df$ngram.size,token=df$Var1,freq=df$Freq )
}


word.frequencies<-frequency(1)



kable(data.frame(unclass(summary(word.frequencies$freq))))
bigram.frequencies<-frequency(2)
kable(data.frame(unclass(summary(bigram.frequencies$freq))))
trigram.frequencies<-frequency(3)
kable(data.frame(unclass(summary(trigram.frequencies$freq))))
quatrigram.frequencies<-frequency(4)
kable(data.frame(unclass(summary(quatrigram.frequencies$freq))))


total.frequencies<-rbind(word.frequencies,bigram.frequencies,trigram.frequencies,quatrigram.frequencies) 

order.inexes<-rev(order(total.frequencies$freq))

total.frequencies<-total.frequencies[order.inexes,]

##chunk cleanup
rm(word.frequencies,bigram.frequencies,trigram.frequencies,quatrigram.frequencies,order.inexes)






```


Below, a list with the most frequent words is shown (ngram size=1):

```{r show_most_frequent_words}

table.index<-(total.frequencies$ngram.size==1 & total.frequencies$freq>600)
kable(total.frequencies[table.index,c("token","freq")], row.names = FALSE, caption = "Most frequent words")

```

the most frequent bigrams (ngram size=2):
```{r show_most_frequent_2grams}

table.index<-(total.frequencies$ngram.size==2 & total.frequencies$freq>80)
kable(total.frequencies[table.index,c("token","freq")], row.names = FALSE, caption = "Most frequent bigrams")

```

the most frequent trigrams (ngram size=3):
```{r show_most_frequent_3grams}

table.index<-(total.frequencies$ngram.size==3 & total.frequencies$freq>11)
kable(total.frequencies[table.index,c("token","freq")], row.names = FALSE, caption = "Most frequent trigrams")

```

and the most frequent 4-grams (ngram size=4):
```{r show_most_frequent_4grams}

table.index<-(total.frequencies$ngram.size==4 & total.frequencies$freq>3)
kable(total.frequencies[table.index,c("token","freq")], row.names = FALSE, caption = "Most frequent 4-grams")


```



```{r frequency_distribution_calculation}
#calculos previos


calculate.freq.distribution <- function (ngram.size)
{
  frequencies<-total.frequencies[total.frequencies$ngram.size==ngram.size,"freq"]
  result<-data.frame(table(frequencies))
  names(result)<-c("freq","n")
  result$freq<-as.numeric(result$freq)
  result$ngram.size<-ngram.size
  result
}

frequency.distribution<-rbind(calculate.freq.distribution(1),
                              calculate.freq.distribution(2),
                              calculate.freq.distribution(3),
                              calculate.freq.distribution(4))



```


We show now a plot with the distribution of frequencies:


```{r plot_word_histograms}
ggplot(frequency.distribution[frequency.distribution$ngram==1,], aes(x=freq,y=n)) + geom_col() 
```


It seems that mostly all words appears only once or twice, but it is difficult to read this plot because it is very skewed, We show below a plot in the log scale to see this more clearly: 



```{r plot_word_histograms_log}
plot.data<-frequency.distribution[frequency.distribution$ngram==1,]
ggplot(plot.data, aes(x=freq,y=log(n))) + geom_col() 

```


We show now the same plot for the all the others ngram sizes (bigrams, 3-gram, 4-gram)
```{r plot_ngrams_histograms_log}

#Distribution of ngrams
plot.data<-frequency.distribution[frequency.distribution$ngram!=1, ]
ggplot(plot.data, aes(x=freq,y=log(n))) + geom_col() + facet_grid(.~ngram.size) 


```



## Coverage

We will look now about the text coverage. How many unique words cover 50% of all words in the text? 


```{r coverage_calculation}

# We can calculate the coverage iterating the frequency table in sequential order because it is ordered

calculateCoverage <-function (coverage, ngram.size)
{
  freq<-total.frequencies[total.frequencies$ngram.size==ngram.size,"freq"]
  coverage.limit<-sum(freq)*coverage
  for (i in 1:length(freq))
  {
    current.coverage <- sum(freq[1:i])
    if (current.coverage>=coverage.limit)
    {
      return (i)
    }
    
  }
  return (-1)
}


coverages<-data.frame(ngram.size=unique(total.frequencies$ngram.size))

coverages$total.words<-sapply(X=coverages$ngram.size, FUN=function(n) {sum(total.frequencies[total.frequencies$ngram.size==n,]$freq)} )

coverages$total.different.tokens<-sapply(X=coverages$ngram.size, FUN=function(n) {nrow(total.frequencies[total.frequencies$ngram.size==n,])} )
coverages$coverage50<-sapply(X=coverages$ngram.size, FUN=function(n) {calculateCoverage(.5,n)} )
coverages$coverage75<-sapply(X=coverages$ngram.size, FUN=function(n) {calculateCoverage(.75,n)} )
coverages$coverage90<-sapply(X=coverages$ngram.size, FUN=function(n) {calculateCoverage(.9,n)} )

```


For individual words (not ngrams), we've found that 'r coverages$coverage50[1]' words cover 50% of the words in the corpus. 

In the following table, we show the word coverage for 50% of the text, 75% and 90%, for single words, bigram, trigrams and 4-grams.


```{r show_coverages}
kable(coverages, 
  col.names = c("ngram size", "total words", "total distinct words", "50% coverage words", "75% coverage words", "90% coverage words") ,
  caption = "token coverage by ngram size")



```


# Conclusions & future directions

From the frequency distributions we can say that: 

- very few words have very high frequency 
- almost all word have very low frequency 

This is also true for ngrams, but at a lesser scale. 

The provisional idea of the prediction model is:

- try to predict the next word using 4-grams
- if no good prediction is found, try to predict the next word using 3-grams
- if no good prediction is found, try to predict the next word using 2-grams
- if no good prediction is found, try to predict the next word using single words.


The coverage information can be used to store the model parameters efficiently. We have seen that few words cover most of the text, but this do not stand for ngrams, so some data structure will be needed to organize the ngrams according with the words it contains.




# TODO 

- labels in plots y tablas
- check gramatica
- check formato correcto para web
- check orden subsecciones

