---
title: "Prepare training data"
output: html_notebook

---




```{r imports, , include=FALSE}
# The libraries needed to run the code
library("NLP")
library("tm")
library("ggplot2")
library("knitr")

getwd()


```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=3, fig.align="center", cache = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


```



Load data and save in RDS format

```{r, eval=FALSE, include=FALSE}
sampleFileBUENO<-function(myfilename, sample.proportion)
{
  path<-paste0("./raw_data/final/en_US/",myfilename)
  out.path<-paste0("./clean_data/",strsplit(myfilename,split = ".txt")[[1]],".sampled.txt")
  raw.data = readLines(path,warn = FALSE,skipNul = TRUE);
  sample<-1==rbinom(n=length(raw.data), p=sample.proportion,size=1)
  data<-raw.data[sample]
  # saveRDS(object = data, file = out.path)
  writeLines(data,con=out.path)
  
}


sampleFile<-function(myfilename, sample.no.se.usa)
{
  path<-paste0("./raw_data/final/en_US/",myfilename)
  out.path<-paste0("./clean_data/",strsplit(myfilename,split = ".txt")[[1]],".sampled.txt")
  raw.data = readLines(path,warn = FALSE,skipNul = TRUE,n = 3);
  # saveRDS(object = data, file = out.path)
  writeLines(raw.data,con=out.path)
  
}



 sampleFileBUENO("en_US.blogs.txt", .01)
 sampleFileBUENO("en_US.news.txt", .01)
 sampleFileBUENO("en_US.twitter.txt", .01)


```




```{r}
corpus.source<-DirSource("./clean_data/")
corpus.source
corpus<-VCorpus(corpus.source, readerControl = list(reader=readPlain,language = "en"))
summary(corpus)


```


```{r}

 corpus <- tm_map(corpus, removeNumbers)
 corpus <- tm_map(corpus, removePunctuation)
 corpus <- tm_map(corpus , stripWhitespace)
# corpus <- tm_map(corpus, tolower)
# corpus <- tm_map(corpus, removeWords, stopwords("english")) # quitar
```



```{r}
#dtm <- DocumentTermMatrix(corpus)
#inspect(dtm)
```



```{r}

#Weka no funciona muy bien paralelizado
#BigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 1, max = 1))
#tdm <- TermDocumentMatrix(corpus, control = list(tokenize = BigramTokenizer))
```



Tokenizin all ngrams in one iteration
```{r}
MultigramTokenizer<-  function(x) {
    words<-words(x)
    bigrams<-ngrams(words(x),2)
    trigrams<-ngrams(words(x),3)

    words<-unlist(lapply(words, paste, collapse = " "), use.names = FALSE)  
    bigrams<-unlist(lapply(bigrams, paste, collapse = " "), use.names = FALSE)  
    trigrams<-unlist(lapply(trigrams, paste, collapse = " "), use.names = FALSE)  

    
    c(words,bigrams, trigrams)
    
}


```

Tokenization (all ngrams)
```{r}


tmd<-TermDocumentMatrix(corpus, control=list(tokenize=MultigramTokenizer))

tmd
```


Training data construction


```{r}
#totalTMD<-rbind(unigramMatrix,bigramMatrix, trigramMatrix)
#attributes(totalTMD) <- attributes(bigramMatrix)
total.frequencies<-apply(as.matrix(tmd), FUN=sum, MARGIN=1) # Add for all documents
# El term es el índice

get.gram.head <- function (s)
{
  paste(s[1:length(s)-1],collapse = " ")
}


get.gram.tail <- function (s)
{
  s[length(s)]
}


test <- data.frame(
  term=names(total.frequencies),
  freq=total.frequencies,
  gram.size=sapply(strsplit(names(total.frequencies)," "), FUN =length),
  gram.head=sapply(strsplit(names(total.frequencies)," "), FUN =get.gram.head),
  gram.tail=sapply(strsplit(names(total.frequencies)," "), FUN =get.gram.tail)
)

test

```




- tm::c.VCorpus		Combine Corpora, Documents, Term-Document Matrices, and Term Frequency Vectors
- tm::findAssocs		Find Associations in a Term-Document Matrix
- tm::findMostFreqTerms		Find Most Frequent Terms
- tm::inspect		Inspect Objects
- tm::TermDocumentMatrix		Term-Document Matrix
- tm::plot.TermDocumentMatrix		Visualize a Term-Document Matrix
- tm::tm_term_score		Compute Score for Matching Terms
- tm::plot.TermDocumentMatrix		Visualize a Term-Document Matrix


